A Constraint Solving Approach to Parikh Images of Regular Languages

### 2022-07-08 main.tex
updated proofs

Since \cref{tbl:rules:multi} only extends the existing rules of
\cref{tbl:rules:single}, we only have to prove termination and solution
preservation for the new rules.

\subsubsection{\Calculus{} for products of automata terminates}
\begin{lemma}\label{lma:multi-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the product
  version of $\Image$ only occurs positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
\end{lemma}

\begin{proof}
  The rule~\ExpandM{} can only be applied finitely often since its side condition only allows applying it once per instance of an $\Image$ predicate containing a product, and the only rule that introduces a new $\Image$ predicate with a product, $\Materialise$ does so with non-overlapping automata from the $\Image$ predicate it removes. Therefore, the number of automata in $\Image$ predicates is strictly decreasing.

  The rule \Materialise{} can similarly only be used finitely many times, as
  each application reduces the number of automata in the product of $\Image$ by
  one automaton, until only one remains and \cref{lma:single-terminates} for
  single-automaton instances applies.
\end{proof}

\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}

\begin{lemma}\label{lma:multi-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:multi}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
  conclusion~$\SomeClause$ if and only if it satisfies one of the
  premises~$\SomeClause_i$.
\end{lemma}

  \ExpandM{} applies \Expand{} to every automata $\Automaton_i$ of the product. 
  \begin{itemize}
  \item Assume that $\beta$ satisfies the conclusion, which means that
    there is some accepting path in the product $\Automaton_1 \times \ldots \times \Automaton_k$ such that
    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton_1 \times \ldots \times \Automaton_k)}$ with
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$ and $\val_\beta(\MonoidElement) =
    \Map(\WordOf(\Path))$. That immediately implies that every automaton $\Automaton_i$ would satisfy its $\Image{}$ constraints individually, and hence also the constraints generated by \ExpandM{}. \Fudge{Missing Materialise!}    
  \item Assume that $\beta$ satisfies the premise, which implies that
    $\val_\beta(\Filter)$ describes a consistent, connected flow of
    the automaton. By the same argument as in
    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
    that each transition~$\Transition$ occurs on $\Path$ exactly
    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
  \end{itemize}

  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
  then a goal must contain~$\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
  formulas in $\SomeInequalities$ are inconsistent, an application of
  \Subsume{} is trivially solution-preserving; therefore assume that
  $\SomeInequalities$ is consistent, which means that it contains
  exactly one of $\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for each $\Transition$.
\end{proof}

>>>>

Since \cref{tbl:rules:multi} only extends the existing rules of
\cref{tbl:rules:single}, we focus on the differences compared
to the calculus for a single automaton.

%\subsubsection{\Calculus{} for products of automata terminates}
\begin{lemma}\label{lma:multi-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the product
  version of $\Image$ only occurs positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
\end{lemma}

\begin{proof}
  The rule \Materialise{} can similarly only be used finitely many times, as
  each application reduces the number of automata in the product of $\Image$ by
  one automaton, until only one remains and \cref{lma:single-terminates} for
  single-automaton instances applies.

  This implies that also the rule~\ExpandM{} can only be applied
  finitely often since its side condition only allows applying it once
  per instance of an $\Image$ predicate containing a product, and only finitely
  many instances of $\Image$ can be introduced on each branch.
\end{proof}

%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}

Since our calculus now includes a rule introducing new variables, the
\Materialise{} rule, we have to slightly generalise the notion of
solution-preservation:
%
\begin{lemma}\label{lma:multi-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:multi}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
  there is an extension~$\beta' of \beta$ satisfying one of the
  premises~$\SomeClause_i$.
\end{lemma}

\begin{proof}
  We have to consider the two new rules in \cref{tbl:rules:multi}. The
  result is immediate for \ExpandM{}, since this rule does not remove the
  $\Image$ predicate from a proof goal, and the newly introduced formulas
  are all implied by the $\Image$ predicate.

  For \Materialise{}, observe that the existence of an accepting path
  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
  same word. The path in the product will satisfy the flow equations
  and connectedness, and it will be related to the individual paths as
  stipulated by the \BindingSum{} predicate.
\end{proof}
